---
title: 'Graphs for DRI Needs Assessment report'
author: 'Jeremy Buhler'
date: 'Dec 10, 2021'
output: html_document
---

## Load required packages
assuming they're already installed
  
```{r}
library(dplyr)
library(tibble)
library(tidyr)
library(tidyselect)
library(ggplot2)
library(stringr)
library(forcats)
library(RColorBrewer)
```


## Import source files

  - DRI_survey_results_20210927.csv __(output as downloaded from Qualtrics with default export settings)__
  - DRI_survey_spam_ids.csv __(list of ResponseIDs previously identified as spam)__

```{r}
responses <- read.csv('../DRI_survey_results_20210927.csv')
spam <- read.csv('../DRI_survey_spam_ids.csv')
```

## Remove first two non-header rows
Qualtrics exports include three header rows; we only need the first *CAUTION: run only once; each time it runs the first two records of the `responses` dataframe are deleted.*

```{r}
responses <- slice(responses, -1:-2)
```

## Limit to responses where "finished" field = TRUE
Qualtrics records partial responses where the respondent never clicked submit; delete these

```{r}
responses <- responses %>%
  filter(Finished == "TRUE")
```

## Remove records identified as spam

```{r}
spam_removed <- anti_join(responses, spam, by="ResponseId" )
```

## Create "discipline" variable that recodes "other" values into best match from survey options. ('x' value recoded to the group with the largest number of responses.)
```{r}
spam_removed <- spam_removed %>%
  mutate(discipline = case_when(
    Q6_8_TEXT == "Computer Science" ~ "Engineering and Technology",
    Q6_8_TEXT == "Creative Technologies - Visual Art" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "Mathematics" ~ "Physical Sciences",
    Q6_8_TEXT == "Math" ~ "Physical Sciences",
    Q6_8_TEXT == "Human Kinetics" ~ "Medicine and Health Sciences",
    Q6_8_TEXT == "Education" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "Law" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "Social Work " ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "interdisciplinary environmental studies" ~ "Ecology and Environmental Sciences",
    Q6_8_TEXT == "x" ~ "Medicine and Health Sciences",
    Q6_8_TEXT == "Education" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "Commerce" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "Applied Science" ~ "Engineering and Technology",
    Q6_8_TEXT == "Pharmaceutical Scs" ~ "Medicine and Health Sciences",
    Q6_8_TEXT == "Architecture" ~ "Engineering and Technology",
    Q6_8_TEXT == "Architecture " ~ "Engineering and Technology",
    Q6_8_TEXT == "Music" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "Education " ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "Architecture" ~ "Engineering and Technology",
    Q6_8_TEXT == "Business School" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "Business" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "health and exercise sciences" ~ "Medicine and Health Sciences",
    Q6_8_TEXT == "Business Administration" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "neuroscience (cognitive)" ~ "Medicine and Health Sciences",
    Q6_8_TEXT == "Business School" ~ "Social Sciences and Humanities",
    Q6_8_TEXT == "Architecture" ~ "Engineering and Technology",
    Q6_8_TEXT == "statistical/data science" ~ "Physical Sciences",
    TRUE ~ Q6
  ))
```


## Remove variables
Remove variables that aren't required for the graphs:

  - survey metadata (i.e. all variables that don't start with "Q#")
  - personal information (Q29, Q31, etc.)
  - variables with the same value in all responses (e.g. where *no one* provided a response)
  - "please specify" fields used by respondents who select "Other"

```{r}
variables_removed <- spam_removed %>%
  select(matches("^Q\\d"),discipline) %>%
  select(-(one_of("Q29","Q31","Q32_1","Q32_2"))) %>%
  select_if(~!n_distinct(.) < 2) %>%
  select(-matches("TEXT"))
```

## Add identifier variable (unique respondent ID not linked to Qualtrics)

```{r}
variables_removed <- rowid_to_column(variables_removed, var="id")
```


## Pivot from wide to long
The export from Qualtrics has one row per respondent, with columns for each question and for each *option* in "select all that apply" questions. Graphing is easier if the data is normalized with each question on its own row. In this process we also split some questions into two variables. For example, a case where the "SupportWhere_Machine learning" variable has a value of "UBC IT" will look like this:

  | variable name | value            |
  | ---           | ---              |
  | question      | SupportWhere     |
  | item          | Machine learning |
  | response      | UBC IT           |
  
### Step 1: Pivot 
```{r}
pivot <- variables_removed %>%
  pivot_longer(
    cols = matches("_"),
    names_to = "question",
    values_to = "response"
  )
```

### Step 2: Split
```{r}
pivot2 <- pivot %>%
  separate(question, c("question","item"),"_", remove = TRUE)
```

### Step 3: Rename values in new columns
```{r}
pivot3 <- pivot2 %>%
  mutate(question = case_when(
                        question == "Q18"   ~ "Support",
                        question == "Q19.1" ~ "SupportWhere",
                        question == "Q19.2" ~ "SupportSatisfaction",
                        question == "Q17"   ~ "Training",
                        question == "Q22.1" ~ "TrainingWhere",
                        question == "Q22.2" ~ "TrainingSatisfaction"
  )) %>%
  
  # Create new "item_group" variable reflecting visual grouping of items in the survey questionnaire (groups in survey were not labelled). This could potentially be used to break up items in graphs. 
  
  mutate(item_group = case_when(
                        item %in% c("1","2","3","4","5","6","7") ~ 'Group 1',
                        item %in% c("8","9","10","11","12","13","14","15") ~ 'Group 2',
                        item %in% c("16","17","18","19","20","21") ~ 'Group 3',
                        item %in% c("22","23","24","25") ~ 'Group 4',
                        TRUE ~ 'other'
  )) %>%                      
  
  
  mutate (item = case_when(
                        item == '1' ~ 'Project management',
                        item == '2' ~ 'Study/data design',
                        item == '3' ~ 'Data management plan',
                        item == '4' ~ 'Data destruction',
                        item == '5' ~ 'Active data storage',
                        item == '6' ~ 'Data transfer',
                        item == '7' ~ 'Data deposit/preservation',
                        item == '8' ~ 'Coding/programming',
                        item == '9' ~ 'Visualization',
                        item == '10' ~ 'Machine learning',
                        item == '11' ~ 'GIS',
                        item == '12' ~ 'Database/app development',
                        item == '13' ~ 'Digital sustainability',
                        item == '14' ~ 'Survey development',
                        item == '15' ~ 'Analysis & statistics',
                        item == '16' ~ 'Ethics compliance',
                        item == '17' ~ 'Privacy',
                        item == '18' ~ 'Security',
                        item == '19' ~ 'Data/research agreements',
                        item == '20' ~ 'Funding',
                        item == '21' ~ 'Inter-institution collab.',
                        item == '22' ~ 'Hardware purchasing',
                        item == '23' ~ 'Contributed systems',
                        item == '24' ~ 'High perf. computing',
                        item == '25' ~ 'Cloud computing',
                        item == '26' ~ 'Other'
  ))
  
  
  
```


## Create dataframes for graphs

 - create `short_response` field for legends
 - remove `other` fields
 - Identify set of responses that items in graph should be sorted by (these are assigned value of "1" in the `sort_include` variable)

```{r}
graph_data <- pivot3 %>%
  mutate(short_response = str_match(response, "(.*)(support|training)")[,1]) %>%
  drop_na(short_response) %>%
  filter(item != "Other") %>%
  mutate(sort_include = if_else(str_starts(short_response,"I would have liked")|str_starts(short_response, "I sought additional"),1,0))
```

# Demographics graphs
```{r}

demo_data <- pivot3 %>%
  select(id, Q5, discipline, Q7) %>%
  mutate(Q7 = str_remove(Q7,"..please specify.")) %>%
  distinct()

demo_data %>%
    ggplot(
    aes(y=fct_rev(fct_infreq(discipline))))+
    geom_bar(fill="#4EB3D3") +
    labs(y="",x="# of respondents", title="Research discipline")+
    scale_fill_brewer(palette = "Blues") +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10, color="#606060"), plot.caption=element_text(face = "italic", color="#606060",margin=margin(15,0,0,0))) 

ggsave("graphs/overall_discipline.png", width=6, height=2.5, units="in")

demo_data %>%
    ggplot(
    aes(y=fct_rev(fct_infreq(Q5))))+
    geom_bar(fill="#4EB3D3") +
    labs(y="",x="# of respondents", title="Position at UBC")+
    scale_fill_brewer(palette = "Blues") +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10, color="#606060"), plot.caption=element_text(face = "italic", color="#606060",margin=margin(15,0,0,0))) 

ggsave("graphs/overall_position.png", width=6, height=2.5, units="in")

demo_data %>%
    ggplot(
    aes(y=fct_rev(fct_infreq(Q7))))+
    geom_bar(fill="#4EB3D3") +
    labs(y="",x="# of respondents", title="Primary work location")+
    scale_fill_brewer(palette = "Blues") +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10, color="#606060"), plot.caption=element_text(face = "italic", color="#606060",margin=margin(15,0,0,0))) 
  
ggsave("graphs/overall_location.png", width=6, height=1.5, units="in")
```

 
# Overall "need for training/support" graph
```{r}
for (i in type) {
  
graph_data_subset <- graph_data %>%
  filter(question == i)

# `item` ordered by totals of `sort_include` variable
graph_data_subset$item <- fct_reorder(graph_data_subset$item,graph_data_subset$sort_include,sum)

overall_graph <- graph_data_subset %>%
  mutate(response = str_match(response,"(.*(additional|liked))")[,2]) %>%
  mutate(response = str_replace(response,"liked","liked*")) %>%
  group_by(item_group,item,short_response) %>%
  ggplot(
    aes(y=item, fill=response))+
    geom_bar() +
    labs(y="",x="# of respondents", fill="", title=paste("Need for",tolower(i),"by topic"), caption=paste("* I would have liked",tolower(i),"but did not know it was available"))+
      scale_fill_brewer(palette = "Greens", limits = c("I would have liked*", "I sought additional", "I did not need additional"), direction=-1) +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10, color="#606060"), plot.caption=element_text(face = "italic", color="#606060",margin=margin(15,0,0,0))) 

ggsave(paste0("graphs/detailed_",tolower(i),".png"), width=7.29, height=5.5, units="in")
print(overall_graph)
}

```


# Units for which individual graphs are required
```{r}
unit <- c("ARC", "Library", "UBC IT")
```

# Type of graph (support/training)
```{r}
type <- c("Support", "Training")
```


```{r}
for (i in type) {
  
graph_w <- pivot3 %>%
  select(id,question,item,response) %>%
  filter(response != "") %>%
  filter(question == paste0(i,"Where")) %>%
  separate_rows(response,sep=",") %>%
  mutate(response = case_when(
            response == "UBC Library" ~ "Library",
            response == "UBC ARC" ~ "ARC",
            response == "UBC IT" ~ "UBC IT",
            str_detect(response,"outside UBC") ~ "Outside UBC",
            TRUE ~ "Other UBC")) %>%
  distinct()

for (u in unit) {
graph_w <- graph_w %>%
  mutate(response = fct_relevel(response, 
            u, after=0)) %>%
  mutate(response = fct_relevel(response, "Other UBC", after=Inf)) %>%
  mutate(response = fct_relevel(response, "Outside UBC", after=Inf)) %>%
  mutate(response = fct_rev(response)) %>%
  mutate(sort_include = if_else(response == u,1,0)) %>%
  mutate(item = fct_reorder(item,sort_include,sum))

output_w <- graph_w %>% 
  ggplot(
    aes(y=item, fill=response))+
    geom_bar() +
    labs(y="",x="# of respondents", fill="", title=paste("Places",tolower(i), "is sought, by topic:",u,"view"))+
    scale_fill_manual(values=c('#FEE6CE','#FDBB83','#F16813','#A63602','#1F78B4')) +
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10, color="#606060"), plot.caption=element_text(face = "italic", color="#606060",margin=margin(15,0,0,0)))

ggsave(paste0("graphs/",u,"_",tolower(i),"_where.png"), width=7.29, height=5.5, units="in")
print(output_w)
  
}
}
```


### Support/training overall graph (# of instances)

```{r}
# set custom palette so colors match other graphs in the report
custom_greens <- brewer.pal(3,"Greens")

for (i in type) {

output_overall <- graph_data %>% 
  filter(question == i) %>%
  mutate(short_response = str_remove(short_response," (training|support)")) %>%
  mutate(short_response = str_replace(short_response,"liked","liked*")) %>%
  ggplot(
    aes(y=short_response, fill=short_response,))+
    geom_bar() +
    labs(y="",x="# of instances", fill="", title=paste("Need for",tolower(i)), caption = paste('* I would have liked',tolower(i), "but did not know it was available") )+
    scale_fill_manual(values = custom_greens) +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10), aspect.ratio = 1/4, legend.position = "none", plot.caption=element_text(face = "italic", color="#606060",margin=margin(15,0,0,0)))

ggsave(paste0("graphs/overall_",tolower(i),".png"), width=5.5, height=2, units="in")
print(output_overall)
}
```




## Where support/training is sought
```{r}
# reshape data
unit_mentions <- pivot3 %>%
  select(id,question,item,response) %>%
  filter(question == "SupportWhere" | question == "TrainingWhere", response != "")

unit_mentions$response <- as.character(unit_mentions$response)

unit_mentions <- unit_mentions %>%
separate_rows(response, sep = ",")

# set custom palette
custom_yellow <- brewer.pal(4,"YlOrBr")[2:3]

# plot training/support graph
unit_mentions %>%
  mutate(color_key = if_else((response == "Support outside UBC"|response == "Training outside UBC"),"1","0")) %>%
  mutate(response = str_trim(str_remove(response,"([Ss]upport|[Tt]raining) ?"))) %>%
  mutate(response = str_replace(response,"out","Out")) %>%
  mutate(question = str_remove(question,"Where")) %>%
  ggplot(
    aes(y=fct_rev(fct_infreq(response)), fill=color_key)) +
    labs(y="",x="# of mentions", title="Places support and training are sought", subtitle="Respondents could select more than one place for each topic")+
    geom_bar() +
    scale_fill_manual(values = custom_yellow) +
    facet_wrap(vars(question)) +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10, color="#606060"), plot.caption=element_text(face = "italic", color="#606060"), legend.position = "none")

ggsave("graphs/overall_where.png", width=7.29, height=3.5, units="in", dpi=320)
```

