---
title: 'Graphs for DRI Needs Assessment report'
author: 'Jeremy Buhler'
date: 'Nov 9, 2021'
output: html_document
---

## Install/load packages
Check for required packages, install if needed

```{r}
# list.of.packages <- c("dplyr","tibble","tidyr","tidyselect","ggplot2","stringr","forcats","RColorBrewer")
# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
# if(length(new.packages)) install.packages(new.packages)
```

Load required packages
  
```{r}
library(dplyr)
library(tibble)
library(tidyr)
library(tidyselect)
library(ggplot2)
library(stringr)
library(forcats)
library(RColorBrewer)
```


## Import source files

  - DRI_survey_results_20210927.csv __(output as downloaded from Qualtrics with default export settings)__
  - DRI_survey_spam_ids.csv __(list of ResponseIDs previously identified as spam)__

```{r}
responses <- read.csv('../DRI_survey_results_20210927.csv')
spam <- read.csv('../DRI_survey_spam_ids.csv')
```

## Remove first two non-header rows
Qualtrics exports include three header rows; we only need the first *CAUTION: run only once; each time it runs the first two records of the `responses` dataframe are deleted.*

```{r}
responses <- slice(responses, -1:-2)
```

## Limit to responses where "finished" field = TRUE
Qualtrics records partial responses where the respondent never clicked submit; delete these

```{r}
responses <- responses %>%
  filter(Finished == "TRUE")
```

## Remove records identified as spam

```{r}
spam_removed <- anti_join(responses, spam, by="ResponseId" )
```

## Remove variables
Remove variables that aren't required for the graphs:

  - survey metadata (i.e. all variables that don't start with "Q#")
  - personal information (Q29, Q31, etc.)
  - variables with the same value in all responses (e.g. where *no one* provided a response)
  - "please specify" fields used by respondents who select "Other"

```{r}
variables_removed <- spam_removed %>%
  select(matches("^Q\\d")) %>%
  select(-(one_of("Q29","Q31","Q32_1","Q32_2"))) %>%
  select_if(~!n_distinct(.) < 2) %>%
  select(-matches("TEXT"))
```

## Add identifier variable (unique respondent ID not linked to Qualtrics)

```{r}
variables_removed <- rowid_to_column(variables_removed, var="id")
```

## Pivot from wide to long
The export from Qualtrics has one row per respondent, with columns for each question and for each *option* in "select all that apply" questions. Graphing is easier if the data is normalized with each question on its own row. In this process we also split some questions into two variables. For example, a case where the "SupportWhere_Machine learning" variable has a value of "UBC IT" will look like this:

  | variable name | value            |
  | ---           | ---              |
  | question      | SupportWhere     |
  | item          | Machine learning |
  | response      | UBC IT           |
  
### Step 1: Pivot 
```{r}
pivot <- variables_removed %>%
  pivot_longer(
    cols = matches("_"),
    names_to = "question",
    values_to = "response"
  )
```

### Step 2: Split
```{r}
pivot2 <- pivot %>%
  separate(question, c("question","item"),"_", remove = TRUE)
```

### Step 3: Rename values in new columns
```{r}
pivot3 <- pivot2 %>%
  mutate(question = case_when(
                        question == "Q18"   ~ "Support",
                        question == "Q19.1" ~ "SupportWhere",
                        question == "Q19.2" ~ "SupportSatisfaction",
                        question == "Q17"   ~ "Training",
                        question == "Q22.1" ~ "TrainingWhere",
                        question == "Q22.2" ~ "TrainingSatisfaction"
  )) %>%
  
  mutate (item = case_when(
                        item == '1' ~ 'Project management',
                        item == '2' ~ 'Study/data design',
                        item == '3' ~ 'Data management plan',
                        item == '4' ~ 'Data destruction',
                        item == '5' ~ 'Active data storage',
                        item == '6' ~ 'Data transfer',
                        item == '7' ~ 'Data deposit/preservation',
                        item == '8' ~ 'Coding/programming',
                        item == '9' ~ 'Visualization',
                        item == '10' ~ 'Machine learning',
                        item == '11' ~ 'GIS',
                        item == '12' ~ 'Database/app development',
                        item == '13' ~ 'Digital sustainability',
                        item == '14' ~ 'Survey development',
                        item == '15' ~ 'Analysis & statistics',
                        item == '16' ~ 'Ethics compliance',
                        item == '17' ~ 'Privacy',
                        item == '18' ~ 'Security',
                        item == '19' ~ 'Data/research agreements',
                        item == '20' ~ 'Funding',
                        item == '21' ~ 'Inter-institution collab.',
                        item == '22' ~ 'Hardware purchasing',
                        item == '23' ~ 'Contributed systems',
                        item == '24' ~ 'High perf. computing',
                        item == '25' ~ 'Cloud computing',
                        item == '26' ~ 'Other'
  ))
```


## Create dataframes for graphs

 - create `short_response` field for legends
 - remove `other` fields
 - Identify set of responses that items in graph should be sorted by (these are assigned value of "1" in the `sort_include` variable)

```{r}
graph_data <- pivot3 %>%
  mutate(short_response = str_match(response, "(.*)(support|training)")[,1]) %>%
  drop_na(short_response) %>%
  filter(item != "Other") %>%
  mutate(sort_include = if_else(str_starts(short_response,"I would have liked")|str_starts(short_response, "I sought additional"),1,0))
```

# Overall "need for training/support" graph
```{r}
overall_graph <- graph_data %>%
  filter(question == "Support"|question == "Training") %>%
  mutate(response = str_match(response,"(.*(additional|liked))")[,2]) %>%
  group_by(item,short_response) %>%
  drop_na(short_response) %>%
  ggplot(
    aes(y=item, fill=response))+
    geom_bar() +
    labs(y="",x="# of respondents", fill="", title="Support and training by topic: all topics", caption="All topics selected by at least one survey respondent")+
    scale_fill_brewer(palette = "Greens")+
    facet_wrap(vars(question)) +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10), plot.caption=element_text(face = "italic", color="#606060",margin=margin(15,0,0,0)))

# ggsave(paste0("graphs/",str_replace(i," ","_"),"_support-training.png"))
print(overall_graph)

```


# Units for which graphs are required
```{r}
unit <- c("UBC ARC", "UBC Library", "UBC IT", "UBC ORS")
```


# Unit-specific "need for training/support" graphs
```{r}
for (i in unit) {

# create subset for SupportWhere graphs
graph_w <- pivot3 %>%
  filter(question == "SupportWhere"|question == "TrainingWhere",item != "Other", response != "") %>%
  mutate(short_response = case_when(
       str_detect(response, paste0(".*",i,".*")) ~ i,
       str_detect(response,"Support outside UBC", negate = TRUE) ~ "Other UBC unit/dept",
       TRUE ~ "Outside UBC")) %>%
    mutate(sort_include = 1)

# reorder levels so selected "unit" is always the same color (otherwise color is assigned alphabetically)
graph_w$short_response <- factor(graph_w$short_response, levels = c("Outside UBC","Other UBC unit/dept",i))

# create list of `items` where at least one person sought support from selected unit 
unit_subset <- graph_w %>%
  filter(short_response == i) %>%
  select(item) %>%
  distinct()
# `item` ordered by totals of `sort_include` variable
graph_data$item <- fct_reorder(graph_data$item,graph_data$sort_include,sum)

# generate support graphs
output_st <- graph_data %>%
  filter(question == "Support"|question == "Training", item %in% unit_subset$item) %>%
  mutate(response = str_match(response,"(.*(additional|liked))")[,2]) %>%
  group_by(item,short_response) %>%
  drop_na(short_response) %>%
  ggplot(
    aes(y=item, fill=response))+
    geom_bar() +
    labs(y="",x="# of respondents", fill="", title=paste("Support and training by topic:",i,"subset"), caption=paste("Topics for which 1 or more respondents sought support or training from",i))+
    scale_fill_brewer(palette = "Greens")+
    facet_wrap(vars(question)) +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10), plot.caption=element_text(face = "italic", color="#606060",margin=margin(15,0,0,0)))

# ggsave(paste0("graphs/",str_replace(i," ","_"),"_support-training.png"))
print(output_st)
}
```


# Unit-specific "where sought support/training" graphs
```{r}
for (i in unit) {

# create subset for TrainingWhere graphs
graph_w <- pivot3 %>%
  filter(question == "TrainingWhere"|question == "SupportWhere", item != "Other", response != "") %>%
  mutate(short_response = case_when(
       str_detect(response, paste0(".*",i,".*")) ~ i,
       str_detect(response,"outside UBC", negate = TRUE) ~ "Other UBC unit/dept",
       TRUE ~ "Outside UBC")) %>%
    mutate(sort_include = 1)

# reorder levels so selected "unit" is always the same color (otherwise color is assigned alphabetically)
graph_w$short_response <- factor(graph_w$short_response, levels = c("Outside UBC","Other UBC unit/dept",i))

# create list of `items` where at least one person sought support from selected unit 
unit_subset <- graph_w %>%
  filter(short_response == i) %>%
  select(item) %>%
  distinct()

# generate trainingWhere/supportWhere graphs
output_w <- graph_w %>% 
  filter(short_response != "NA/blank") %>%
  filter(item %in% unit_subset$item) %>%
  group_by(item,short_response) %>%
  ggplot(
    aes(y=fct_rev(fct_infreq(item)), fill=short_response))+
    geom_bar() +
    labs(y="",x="# of respondents", fill="", title=paste("Places training is sought, by topic:",i,"subset"), caption=paste("Limited to topics for which 1 or more respondents sought support or training from",i))+
    scale_fill_brewer(palette = "Oranges") +
    facet_wrap(vars(question)) +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10), plot.caption=element_text(face = "italic", color="#606060",margin=margin(15,0,0,0)))

ggsave(paste0("graphs/",str_replace(i," ","_"),"_trainingWhere.png"))
print(output_w)
}
```



### Support/training overall graph (# of instances)

```{r}
# set custom palette so colors match other graphs in the report
custom_greens <- brewer.pal(3,"Greens")[2:3]

graph_data %>% 
  filter(str_detect(short_response,"I would have liked") == TRUE | str_detect(short_response,"I sought additional") == TRUE) %>%
  mutate(short_response = str_remove(short_response," (training|support)")) %>%
  ggplot(
    aes(y=short_response, fill=short_response,))+
    geom_bar() +
    facet_wrap(vars(question)) +
    labs(y="",x="# of instances", fill="", title="Need for support/training", caption = '* "I would have liked support/training but did not know it was available"' )+
    scale_fill_manual(values = custom_greens) +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10), aspect.ratio = 1/4, legend.position = "none", plot.caption=element_text(size=9,color='#808080',margin=margin(15,0,0,0)))

ggsave("graphs/overall_support-training.png")
```




## Where support/training is sought
```{r fig.height=3}
# reshape data
unit_mentions <- pivot3 %>%
  select(id,question,item,response) %>%
  filter(question == "SupportWhere" | question == "TrainingWhere", response != "")

unit_mentions$response <- as.character(unit_mentions$response)

unit_mentions <- unit_mentions %>%
separate_rows(response, sep = ",")

# set custom palette
custom_yellow <- brewer.pal(4,"YlOrBr")[2:3]

# plot training/support graph
unit_mentions %>%
  mutate(color_key = if_else((response == "Support outside UBC"|response == "Training outside UBC"),"1","0")) %>%
  mutate(response = str_trim(str_remove(response,"([Ss]upport|[Tt]raining) ?"))) %>%
  mutate(response = str_replace(response,"out","Out")) %>%
  mutate(question = str_remove(question,"Where")) %>%
  ggplot(
    aes(y=fct_rev(fct_infreq(response)), fill=color_key)) +
    labs(y="",x="# of mentions", title="Places support/training is sought")+
    geom_bar() +
    scale_fill_manual(values = custom_yellow) +
    facet_wrap(vars(question)) +
    theme_minimal() + theme(panel.grid.major.y = element_line(color = "white"), plot.subtitle=element_text(size=10), plot.caption=element_text(face = "italic", color="#606060"), legend.position = "none")

ggsave("graphs/overall_where_support-training.png")
```

